The zoon R package for reproducible and shareable species distribution modelling

### 2017-05-16 ms.Rmd
add data accessibility section
switch to word doc

---
title: "The zoon R package for reproducible and shareable species distribution modelling"
author: 
  - Nick Golding$^1$, Tom A. August$^2$, Tim C.D. Lucas$^3$, David J. Gavaghan$^4$, E. Emiel van Loon$^5$ & Greg McInerny$^6$
  - $^1$ School of BioSciences, University of Melbourne, Parkville VIC, Australia
  - $^2$ NERC Centre for Ecology & Hydrology, Crowmarsh Gifford, Maclean Building, Wallingford, OX10 8BB, UK
  - $^3$ Oxford Big Data Institute, Li Ka Shing Centre for Health Information and Discovery Nuffield Department of Medicine, University of Oxford, Oxford, UK
  - $^4$ Department of Computer Science, University of Oxford, Oxford, UK
  - $^5$ Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, NL
  - $^6$ Centre for Interdisciplinary Methodologies, Social Sciences, University of Warwick, Coventry, CV4 7AL, UK
bibliography: zoon_app.bib
csl: mee.csl
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: latex_preamble.tex
    highlight: null
---

@@

```{r flow_diagram, echo = FALSE, out.width = "375px", fig.cap = "The modular SDM structure encoded by a zoon workflow. A) Description of the five module types. B) Flow diagram illustrating how objects are passed between different module types: `data frame' - an dataframe of occurrence records; `raster' - a RasterStack object of the covariates; `model' - a ZoonModel object, generating standardised predictions from a given model. C) The flow diagram implied by chaining two `process' modules. D) The flow diagram implied by listing three `model' modules. Full details of module inputs and outputs, and the effects of listing and chaining each module type are given in the zoon vignette `Building a module'."}
knitr::include_graphics("figs/diagrams.png")
```

@@

\blandscape
```{r feng_papes_plots, cache = FALSE, echo = FALSE, out.width = "450px", fig.cap = "Outputs of the workflow objects `FengPapes' and `FengPapesUpdate'. A) Map of the MaxEnt predicted distribution, with a 5\\% omission rate threshold, by the `PrintMap' module in the workflow `FengPapes' which encodes the core of a published analysis. B) A response curve produced by the `ResponseCurve' module for the first covariate, bio1 in the workflow `FengPapesUpdate', which modifies the original analysis workflow. C) A screenshot of the interactive map produced by the `InteractiveMap' modules in the workflow `FengPapesUpdate', displaying raw occurrence data and predicted distribution over a global map, allowing users to interactively explore their results. White areas are masked due to being in the MESS mask. Any SDM analysis distributed as a zoon workflow can be easily be explored and scrutinized by modifying its output modules using the function `ChangeWorkflow'."}
knitr::include_graphics("figs/combinefengupdate-1.png")
```
\elandscape

@@

```{r maxent_plots, cache = FALSE, echo = FALSE, out.width = "450px", fig.cap = "Prediction of the distribution of the Carolina wren from 6 different models, produced by the workflow `MaxEntComparison'. A) MaxEnt without threshold features. B) MaxNet with default settings. C) MaxNet without regularisation. D) MaxNet with regularisation but only linear features. E) MaxNet without regularisation and only linear features. F) Logistic regression. MaxNet with full features but no regularisation (C) gave the most local complexity, indicative of overfitting to the data. The models with only linear features (D-F) had a broader distribution, indicative of underfitting. Differences between MaxNet with linear features and no  regularisation (E) and logistic regression (F) are due to the downweighting applied to background data in the former."}
knitr::include_graphics("figs/maxent_plotting-1.png")
```

@@

>>>>

---
title: "The zoon R package for reproducible and shareable species distribution modelling"
author: 
  - Nick Golding$^1$, Tom A. August$^2$, Tim C.D. Lucas$^3$, David J. Gavaghan$^4$, E. Emiel van Loon$^5$ & Greg McInerny$^6$
  - $^1$ School of BioSciences, University of Melbourne, Parkville VIC, Australia
  - $^2$ NERC Centre for Ecology & Hydrology, Crowmarsh Gifford, Maclean Building, Wallingford, OX10 8BB, UK
  - $^3$ Oxford Big Data Institute, Li Ka Shing Centre for Health Information and Discovery Nuffield Department of Medicine, University of Oxford, Oxford, UK
  - $^4$ Department of Computer Science, University of Oxford, Oxford, UK
  - $^5$ Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, NL
  - $^6$ Centre for Interdisciplinary Methodologies, Social Sciences, University of Warwick, Coventry, CV4 7AL, UK
bibliography: zoon_app.bib
csl: mee.csl
output:
  word_document:
    fig_caption: true
    highlight: null
---

@@

@@

@@

@@

### Data accessibility

Nine-banded armadillo data were downloaded from GBIF, Carolina Wren data from @maxlike were accessed via the `maxlike` R package. Both datasets can be accessed using the code presented here.

### Figures

```{r flow_diagram, echo = FALSE, fig.cap = "Figure 1. The modular SDM structure encoded by a zoon workflow. A) Description of the five module types. B) Flow diagram illustrating how objects are passed between different module types: 'data frame' - an dataframe of occurrence records; 'raster' - a RasterStack object of the covariates; 'model' - a ZoonModel object, generating standardised predictions from a given model. C) The flow diagram implied by chaining two `process' modules. D) The flow diagram implied by listing three 'model' modules. Full details of module inputs and outputs, and the effects of listing and chaining each module type are given in the zoon vignette 'Building a module'."}
knitr::include_graphics("figs/diagrams.png")
```

```{r feng_papes_plots, cache = FALSE, echo = FALSE, fig.cap = "Figure 2. Outputs of the workflow objects 'FengPapes' and 'FengPapesUpdate'. A) Map of the MaxEnt predicted distribution, with a 5\\% omission rate threshold, by the 'PrintMap' module in the workflow 'FengPapes' which encodes the core of a published analysis. B) A response curve produced by the 'ResponseCurve' module for the first covariate, bio1 in the workflow 'FengPapesUpdate', which modifies the original analysis workflow. C) A screenshot of the interactive map produced by the 'InteractiveMap' modules in the workflow 'FengPapesUpdate', displaying raw occurrence data and predicted distribution over a global map, allowing users to interactively explore their results. White areas are masked due to being in the MESS mask. Any SDM analysis distributed as a zoon workflow can be easily be explored and scrutinized by modifying its output modules using the function 'ChangeWorkflow'."}
knitr::include_graphics("figs/combinefengupdate-1.png")
```

```{r maxent_plots, cache = FALSE, echo = FALSE, fig.cap = "Figure 3. Prediction of the distribution of the Carolina wren from 6 different models, produced by the workflow 'MaxEntComparison'. A) MaxEnt without threshold features. B) MaxNet with default settings. C) MaxNet without regularisation. D) MaxNet with regularisation but only linear features. E) MaxNet without regularisation and only linear features. F) Logistic regression. MaxNet with full features but no regularisation (C) gave the most local complexity, indicative of overfitting to the data. The models with only linear features (D-F) had a broader distribution, indicative of underfitting. Differences between MaxNet with linear features and no  regularisation (E) and logistic regression (F) are due to the downweighting applied to background data in the former."}
knitr::include_graphics("figs/maxent_plotting-1.png")
```
